---
title: "BER Services Surveys"
subtitle: "Preliminary Results"
author: Laurie Binge
date: "17 October 2017"
output:
  beamer_presentation:
    theme: "CambridgeUS"
    colortheme: "dolphin"
    fonttheme: "structurebold" 
    toc: true
    slide_level: 2
---

#Published Results

```{r ReadServices, echo=FALSE, results='hide', message=FALSE, warning=FALSE, cache = TRUE}
setwd("C:\\Users\\Laurie\\OneDrive\\Documents\\BING\\BER Confidence Surveys\\Sentiment")
#change the working directory

suppressMessages(library(ggplot2))
suppressMessages(library(plyr))
suppressMessages(library(dplyr))
suppressMessages(library(reshape2))
suppressMessages(library(stargazer))
suppressMessages(library(xtable))
suppressMessages(library(scales))
suppressMessages(library(quantmod))
suppressMessages(library(vars))
suppressMessages(library(tseries))
suppressMessages(library(urca))

##=====================##
## READING IN THE DATA ##
##=====================##
datums <- read.csv("dates.csv", header=TRUE, sep=",",na.strings = "", skipNul = TRUE)
datums$Datum <- as.Date(datums$Datum, format = "%Y/%m/%d")

pub_s <- read.csv("Services_BER_Published.csv", header=TRUE, sep=",",na.strings = "", skipNul = TRUE)
#ref_s <- read.csv("Ref Series_Services.csv", header=TRUE, sep=",",na.strings = "", skipNul = TRUE)

#================================
#SERVICES
BER.S <- read.csv("Services_05Q2-17Q2.csv", header=TRUE, sep=",",na.strings = "", skipNul = TRUE)
colnames(BER.S)[1:6] <- c("region","id","sector","weight","factor","surveyQ")
BER.S$surveyQ <- toupper(BER.S$surveyQ)
BER.S$sector <- factor(BER.S$sector) #could include labels
BER.S$id <- factor(BER.S$id)
BER.S$region <- factor(BER.S$region)

#Clean SurveyQ variable
BER.S$temp <- NULL
for(i in 1:nrow(BER.S)) {
    ifelse(substr(BER.S$surveyQ[i], 1, 1)==9, 
           BER.S$temp[i] <- paste0("19",BER.S$surveyQ[i],sep=""),
           BER.S$temp[i] <- paste0("20",BER.S$surveyQ[i],sep=""))
}
BER.S$surveyQ <- BER.S$temp
BER.S <- BER.S[,-ncol(BER.S)]
BER.S$surveyQ <- factor(BER.S$surveyQ)

#Eclude Latecomers and old quesions
BER.S <- BER.S[BER.S$Latecomer == FALSE | is.na(BER.S$Latecomer),]
BER.S <- BER.S[,1:21]
BER.S <- BER.S[,!grepl("X",colnames(BER.S))]    

#Replace Values
for(i in 7:21) {
    BER.S[,i] <- replace(BER.S[,i], BER.S[,i]==2, 0)
    BER.S[,i] <- replace(BER.S[,i], BER.S[,i]==3,-1)
}
for(i in 18:21) {
    BER.S[,i] <- replace(BER.S[,i], BER.S[,i]==0, 0.5)
    BER.S[,i] <- replace(BER.S[,i], BER.S[,i]==-1, 0)
}

#----------------------------
#SERVICES
catering <- c(6000,6001,6020,6030,6011)
transport.s <- c(7020,7010,7070,7090,7080,7060,7000,7040,7100,7120,7110,7050)
realestate <- c(8000,8010,8020)
business <- c(8040,8080,8070,8090,8060,8050,8030,
              8150,8120,8210,8180,8140,8160,8190,8100,8200,8230,8130,8110,8170,8240,8220)
#community <- c(9000,9010,9030,9050,9060,9020,9040)
all_s <- as.numeric(as.character(unique(BER.S$sector)))
streke <- unique(BER.S$region)

```

```{r CalcServices, echo=FALSE, results='hide', message=FALSE, warning=FALSE, cache = TRUE}
##=====================##
##Define Functions
##=====================##

ongeweeg <- function(data,sektor,streek,d) {
    data <- data[data$sector %in% sektor & data$region %in% streek,]
    sector <- aggregate(data[,(match("surveyQ",colnames(data))+1):ncol(data)], 
                        by=list(data$surveyQ), FUN=mean, na.rm=TRUE)
    sector$Obs <- aggregate(data$Q2A, by=list(data$surveyQ), FUN=length)[,2]
    sector <- merge(datums[d,],sector, by.x="Date", by.y="Group.1", all=TRUE)
    sector <- sector[,c(1:2,ncol(sector),3:(ncol(sector)-1))]
    sector[,-1:-3] <- na.approx(sector[,-1:-3]*100)
    return(sector)
}


geweeg <- function(data,sektor,streek,d) {
    weeg <- function(temp) {  #calculate weighted mean for each quarter for all columns
        temp <- cbind(factor=temp$factor,temp$factor*temp[(match("surveyQ",colnames(temp))+1):ncol(temp)])
        temp <- colSums(temp, na.rm=TRUE, dims = 1)/    #calculate the sum(wi*xi)/sum(wi)
            sapply(colnames(temp), function(x) sum(temp$factor[!is.na(temp[colnames(temp) == x])]))
        return(temp) #weight only by those that responded to a specific question
    }
    
    data <- data[data$sector %in% sektor & data$region %in% streek,]
    sector <- data.frame()
    for(kwartaal in levels(data$surveyQ)) {
        sector <- rbind(sector,weeg(data[data$surveyQ==kwartaal,]))
    }
    sector[,1] <- levels(data$surveyQ)
    colnames(sector) <- colnames(data)[(match("surveyQ",colnames(data))):ncol(data)]
    sector$Obs <- aggregate(data$Q2A, by=list(data$surveyQ), FUN=length)[,2]
    sector <- merge(datums[d,],sector, by.x="Date", by.y="surveyQ", all=TRUE)
    sector <- sector[,c(1:2,ncol(sector),3:(ncol(sector)-1))]
    sector[,-1:-3] <- na.approx(sector[,-1:-3]*100)
    return(sector)
}


#===========================================
#SERVICES
Services_w <- geweeg(BER.S,all_s,streke,54:102)
Services <- ongeweeg(BER.S,all_s,streke,54:102)

ServicesC <- ongeweeg(BER.S,catering,streke,54:102)
ServicesT <- ongeweeg(BER.S,transport.s,streke,54:102)
ServicesR <- ongeweeg(BER.S,realestate,streke,54:102)
ServicesB <- ongeweeg(BER.S,business,streke,54:102)

WC.S <- ongeweeg(BER.S,all_s,1,54:102)
GP.S <- ongeweeg(BER.S,all_s,6,54:102)
KZN.S <- ongeweeg(BER.S,all_s,5,54:102)

#Interpolasie
p.inter <- c(4,5,7,9,11,13,15:18)
Services[3,p.inter] <- pub_s[3,2:11]
ServicesC[3,p.inter] <- pub_s[3,13:22]
ServicesT[3,p.inter] <- pub_s[3,24:33]
ServicesR[3,p.inter] <- pub_s[3,35:44]
ServicesB[3,p.inter] <- pub_s[3,46:55]

p.inter <- c(4,5,7,9,11,13,15:18)
WC.S[3,p.inter] <- pub_s[3,79:88]
GP.S[3,p.inter] <- pub_s[3,68:77]
KZN.S[3,p.inter] <- pub_s[3,90:99]

```



##Total Services Results
The results for the constraints (Q7-Q10) do not match

```{r figure1, echo=FALSE, message= FALSE, warning=FALSE, cache = TRUE, fig.height=4.5, fig.width=7.5}
#Total Services
indicator_plot <- cbind(Services[,c("Datum","Q1")],pub_s[,"Total_Q1"]) 
colnames(indicator_plot) <- c("Date","Microdata","Published")
indicator_plot <- melt(indicator_plot, id="Date")  # convert to long format
g1 <- ggplot(data=indicator_plot,aes(x=Date, y=value, group=variable, colour=variable)) 
g1 <- g1 + geom_line()
g1 <- g1 + ylab("Indicator") + xlab("")
g1 <- g1 + theme(legend.title=element_blank())
g1 <- g1 + ggtitle("Services: Q1") 
g1 <- g1 + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))
g1 <- g1 + theme(legend.title=element_blank()) 
g1 <- g1 + scale_x_date(labels = date_format("%Y"),breaks = date_breaks("year"))
g1 <- g1 + theme(legend.position="none")

indicator_plot <- cbind(Services[,c("Datum","Q2A")],pub_s[,"Total_Q2A"]) 
colnames(indicator_plot) <- c("Date","Microdata","Published")
indicator_plot <- melt(indicator_plot, id="Date")  # convert to long format
g2 <- ggplot(data=indicator_plot,aes(x=Date, y=value, group=variable, colour=variable)) 
g2 <- g2 + geom_line()
g2 <- g2 + ylab("") + xlab("")
g2 <- g2 + theme(legend.title=element_blank())
g2 <- g2 + ggtitle("Services: Q2A") 
g2 <- g2 + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))
g2 <- g2 + theme(legend.title=element_blank()) 
g2 <- g2 + scale_x_date(labels = date_format("%Y"),breaks = date_breaks("year"))
g2 <- g2 + theme(legend.position="none")

indicator_plot <- cbind(Services[,c("Datum","Q3A")],pub_s[,"Total_Q3A"]) 
colnames(indicator_plot) <- c("Date","Microdata","Published")
indicator_plot <- melt(indicator_plot, id="Date")  # convert to long format
g3 <- ggplot(data=indicator_plot,aes(x=Date, y=value, group=variable, colour=variable)) 
g3 <- g3 + geom_line()
g3 <- g3 + ylab("Indicator") + xlab("")
g3 <- g3 + theme(legend.title=element_blank())
g3 <- g3 + ggtitle("Services: Q3A") 
g3 <- g3 + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))
g3 <- g3 + theme(legend.title=element_blank()) 
g3 <- g3 + scale_x_date(labels = date_format("%Y"),breaks = date_breaks("year"))
g3 <- g3 + theme(legend.position="bottom",plot.margin=unit(c(-0.5,0.4,0,0.4), "cm"))

indicator_plot <- cbind(Services[,c("Datum","Q8")],pub_s[,"Total_Q8"]) 
colnames(indicator_plot) <- c("Date","Microdata","Published")
indicator_plot <- melt(indicator_plot, id="Date")  # convert to long format
g4 <- ggplot(data=indicator_plot,aes(x=Date, y=value, group=variable, colour=variable)) 
g4 <- g4 + geom_line()
g4 <- g4 + ylab("") + xlab("")
g4 <- g4 + theme(legend.title=element_blank())
g4 <- g4 + ggtitle("Services: Q8") 
g4 <- g4 + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))
g4 <- g4 + theme(legend.title=element_blank()) 
g4 <- g4 + theme(legend.position="bottom",plot.margin=unit(c(-0.5,0.4,0,0.4), "cm"))
g4 <- g4 + scale_x_date(labels = date_format("%Y"),breaks = date_breaks("year"))

library(gridExtra)
grid.arrange(g1, g2, g3, g4, ncol=2, nrow =2)
```

##Services Subsector Results: Q1
The results for Q1 do not match up perfectly

```{r figure2, echo=FALSE, warning=FALSE, cache = TRUE, fig.height=4.5, fig.width=7.5}
#Confidence
indicator_plot <- cbind(ServicesC[,c("Datum","Q1")],pub_s[,"Accom_Q1"]) 
colnames(indicator_plot) <- c("Date","Microdata","Published")
indicator_plot <- melt(indicator_plot, id="Date")  # convert to long format
g1 <- ggplot(data=indicator_plot,aes(x=Date, y=value, group=variable, colour=variable)) 
g1 <- g1 + geom_line()
g1 <- g1 + ylab("Indicator") + xlab("")
g1 <- g1 + theme(legend.title=element_blank())
g1 <- g1 + ggtitle("Catering: Q1") 
g1 <- g1 + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))
g1 <- g1 + theme(legend.title=element_blank()) 
g1 <- g1 + scale_x_date(labels = date_format("%Y"),breaks = date_breaks("year"))
g1 <- g1 + theme(legend.position="none")

indicator_plot <- cbind(ServicesR[,c("Datum","Q1")],pub_s[,"RE_Q1"]) 
colnames(indicator_plot) <- c("Date","Microdata","Published")
indicator_plot <- melt(indicator_plot, id="Date")  # convert to long format
g2 <- ggplot(data=indicator_plot,aes(x=Date, y=value, group=variable, colour=variable)) 
g2 <- g2 + geom_line()
g2 <- g2 + ylab("") + xlab("")
g2 <- g2 + theme(legend.title=element_blank())
g2 <- g2 + ggtitle("Real Estate: Q1") 
g2 <- g2 + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))
g2 <- g2 + theme(legend.title=element_blank()) 
g2 <- g2 + scale_x_date(labels = date_format("%Y"),breaks = date_breaks("year"))
g2 <- g2 + theme(legend.position="none")

indicator_plot <- cbind(ServicesT[,c("Datum","Q1")],pub_s[,"Trans_Q1"]) 
colnames(indicator_plot) <- c("Date","Microdata","Published")
indicator_plot <- melt(indicator_plot, id="Date")  # convert to long format
g3 <- ggplot(data=indicator_plot,aes(x=Date, y=value, group=variable, colour=variable)) 
g3 <- g3 + geom_line()
g3 <- g3 + ylab("Indicator") + xlab("")
g3 <- g3 + theme(legend.title=element_blank())
g3 <- g3 + ggtitle("Transport: Q1") 
g3 <- g3 + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))
g3 <- g3 + theme(legend.title=element_blank()) 
g3 <- g3 + scale_x_date(labels = date_format("%Y"),breaks = date_breaks("year"))
g3 <- g3 + theme(legend.position="bottom",plot.margin=unit(c(-0.5,0.4,0,0.4), "cm"))

indicator_plot <- cbind(ServicesB[,c("Datum","Q1")],pub_s[,"BS_Q1"]) 
colnames(indicator_plot) <- c("Date","Microdata","Published")
indicator_plot <- melt(indicator_plot, id="Date")  # convert to long format
g4 <- ggplot(data=indicator_plot,aes(x=Date, y=value, group=variable, colour=variable)) 
g4 <- g4 + geom_line()
g4 <- g4 + ylab("") + xlab("")
g4 <- g4 + theme(legend.title=element_blank())
g4 <- g4 + ggtitle("Business: Q1") 
g4 <- g4 + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))
g4 <- g4 + theme(legend.title=element_blank()) 
g4 <- g4 + theme(legend.position="bottom",plot.margin=unit(c(-0.5,0.4,0,0.4), "cm"))
g4 <- g4 + scale_x_date(labels = date_format("%Y"),breaks = date_breaks("year"))

library(gridExtra)
grid.arrange(g1, g2, g3, g4, ncol=2, nrow =2)
```


##Services Subsector Results: Q2A
The results for Q2A do not match up perfectly

```{r figure3, echo=FALSE, warning=FALSE, cache = TRUE, fig.height=4.5, fig.width=7.5}
#Sales Volumes
indicator_plot <- cbind(ServicesC[,c("Datum","Q2A")],pub_s[,"Accom_Q2A"]) 
colnames(indicator_plot) <- c("Date","Microdata","Published")
indicator_plot <- melt(indicator_plot, id="Date")  # convert to long format
g1 <- ggplot(data=indicator_plot,aes(x=Date, y=value, group=variable, colour=variable)) 
g1 <- g1 + geom_line()
g1 <- g1 + ylab("Indicator") + xlab("")
g1 <- g1 + theme(legend.title=element_blank())
g1 <- g1 + ggtitle("Catering: Q2A") 
g1 <- g1 + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))
g1 <- g1 + theme(legend.title=element_blank()) 
g1 <- g1 + scale_x_date(labels = date_format("%Y"),breaks = date_breaks("year"))
g1 <- g1 + theme(legend.position="none")

indicator_plot <- cbind(ServicesR[,c("Datum","Q2A")],pub_s[,"RE_Q2A"]) 
colnames(indicator_plot) <- c("Date","Microdata","Published")
indicator_plot <- melt(indicator_plot, id="Date")  # convert to long format
g2 <- ggplot(data=indicator_plot,aes(x=Date, y=value, group=variable, colour=variable)) 
g2 <- g2 + geom_line()
g2 <- g2 + ylab("") + xlab("")
g2 <- g2 + theme(legend.title=element_blank())
g2 <- g2 + ggtitle("Real Estate: Q2A") 
g2 <- g2 + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))
g2 <- g2 + theme(legend.title=element_blank()) 
g2 <- g2 + scale_x_date(labels = date_format("%Y"),breaks = date_breaks("year"))
g2 <- g2 + theme(legend.position="none")

indicator_plot <- cbind(ServicesT[,c("Datum","Q2A")],pub_s[,"Trans_Q2A"]) 
colnames(indicator_plot) <- c("Date","Microdata","Published")
indicator_plot <- melt(indicator_plot, id="Date")  # convert to long format
g3 <- ggplot(data=indicator_plot,aes(x=Date, y=value, group=variable, colour=variable)) 
g3 <- g3 + geom_line()
g3 <- g3 + ylab("Indicator") + xlab("")
g3 <- g3 + theme(legend.title=element_blank())
g3 <- g3 + ggtitle("Transport: Q2A") 
g3 <- g3 + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))
g3 <- g3 + theme(legend.title=element_blank()) 
g3 <- g3 + scale_x_date(labels = date_format("%Y"),breaks = date_breaks("year"))
g3 <- g3 + theme(legend.position="bottom",plot.margin=unit(c(-0.5,0.4,0,0.4), "cm"))

indicator_plot <- cbind(ServicesB[,c("Datum","Q2A")],pub_s[,"BS_Q2A"]) 
colnames(indicator_plot) <- c("Date","Microdata","Published")
indicator_plot <- melt(indicator_plot, id="Date")  # convert to long format
g4 <- ggplot(data=indicator_plot,aes(x=Date, y=value, group=variable, colour=variable)) 
g4 <- g4 + geom_line()
g4 <- g4 + ylab("") + xlab("")
g4 <- g4 + theme(legend.title=element_blank())
g4 <- g4 + ggtitle("Business: Q2A") 
g4 <- g4 + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))
g4 <- g4 + theme(legend.title=element_blank()) 
g4 <- g4 + theme(legend.position="bottom",plot.margin=unit(c(-0.5,0.4,0,0.4), "cm"))
g4 <- g4 + scale_x_date(labels = date_format("%Y"),breaks = date_breaks("year"))

library(gridExtra)
grid.arrange(g1, g2, g3, g4, ncol=2, nrow =2)
```


##Regional Services Surveys
```{r figure4, echo=FALSE, warning=FALSE, cache = TRUE, fig.height=4.5, fig.width=7.5}
#Regions
BERplot <- aggregate(BER.S$id, by=list(BER.S$surveyQ,BER.S$region), FUN = length)
BERplot$Group.1 <- as.Date(as.yearqtr(BERplot$Group.1, format = "%YQ%q"))
g <- ggplot(BERplot, aes(x=Group.1, y=x,fill=Group.2))
g <- g + geom_bar(stat="identity")
g <- g + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))
g <- g + scale_fill_discrete(name="Region")
g <- g + scale_y_continuous(labels=comma)
g <- g + scale_x_date(labels = date_format("%Y"),breaks = date_breaks("year"))
g <- g + ylab("Number of Respondents")
g <- g + xlab("Date")
g
```


##Regional Services Results: Q1
The results for Q1 do not match up perfectly

```{r figure5, echo=FALSE, warning=FALSE, cache = TRUE, fig.height=4.5, fig.width=7.5}
indicator_plot <- cbind(WC.S[,c("Datum","Q1")],pub_s[,"WC_Q1"]) 
colnames(indicator_plot) <- c("Date","Microdata","Published")
indicator_plot <- melt(indicator_plot, id="Date")  # convert to long format
g1 <- ggplot(data=indicator_plot,aes(x=Date, y=value, group=variable, colour=variable)) 
g1 <- g1 + geom_line()
g1 <- g1 + ylab("Indicator") + xlab("")
g1 <- g1 + theme(legend.title=element_blank())
g1 <- g1 + ggtitle("WC: Q1") 
g1 <- g1 + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))
g1 <- g1 + theme(legend.title=element_blank()) 
g1 <- g1 + scale_x_date(labels = date_format("%Y"),breaks = date_breaks("year"))
g1 <- g1 + theme(legend.position="none")

indicator_plot <- cbind(GP.S[,c("Datum","Q1")],pub_s[,"GP_Q1"]) 
colnames(indicator_plot) <- c("Date","Microdata","Published")
indicator_plot <- melt(indicator_plot, id="Date")  # convert to long format
g2 <- ggplot(data=indicator_plot,aes(x=Date, y=value, group=variable, colour=variable)) 
g2 <- g2 + geom_line()
g2 <- g2 + ylab("") + xlab("")
g2 <- g2 + theme(legend.title=element_blank())
g2 <- g2 + ggtitle("GP: Q1") 
g2 <- g2 + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))
g2 <- g2 + theme(legend.title=element_blank()) 
g2 <- g2 + scale_x_date(labels = date_format("%Y"),breaks = date_breaks("year"))
g2 <- g2 + theme(legend.position="none")

indicator_plot <- cbind(KZN.S[,c("Datum","Q1")],pub_s[,"KZN_Q1"]) 
colnames(indicator_plot) <- c("Date","Microdata","Published")
indicator_plot <- melt(indicator_plot, id="Date")  # convert to long format
g3 <- ggplot(data=indicator_plot,aes(x=Date, y=value, group=variable, colour=variable)) 
g3 <- g3 + geom_line()
g3 <- g3 + ylab("Indicator") + xlab("")
g3 <- g3 + theme(legend.title=element_blank())
g3 <- g3 + ggtitle("KZN: Q1") 
g3 <- g3 + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))
g3 <- g3 + theme(legend.title=element_blank()) 
g3 <- g3 + scale_x_date(labels = date_format("%Y"),breaks = date_breaks("year"))
g3 <- g3 + theme(legend.position="bottom",plot.margin=unit(c(-0.5,0.4,0,0.4), "cm"))

indicator_plot <- cbind(WC.S[,c("Datum","Q1")],GP.S[,c("Q1")],KZN.S[,c("Q1")]) 
colnames(indicator_plot) <- c("Date","WC","GP","KZN")
indicator_plot <- melt(indicator_plot, id="Date")  # convert to long format
g4 <- ggplot(data=indicator_plot,aes(x=Date, y=value, group=variable, colour=variable)) 
g4 <- g4 + geom_line()
g4 <- g4 + ylab("") + xlab("")
g4 <- g4 + theme(legend.title=element_blank())
g4 <- g4 + ggtitle("Regions: Q1") 
g4 <- g4 + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))
g4 <- g4 + theme(legend.title=element_blank()) 
g4 <- g4 + theme(legend.position="bottom",plot.margin=unit(c(-0.5,0.4,0,0.4), "cm"))
g4 <- g4 + scale_x_date(labels = date_format("%Y"),breaks = date_breaks("year"))

library(gridExtra)
grid.arrange(g1, g2, g3, g4, ncol=2, nrow =2)
```


##Regional Services Results: Q2A
The results for Q2A do not match up perfectly

```{r figure6, echo=FALSE, warning=FALSE, cache = TRUE, fig.height=4.5, fig.width=7.5}
indicator_plot <- cbind(WC.S[,c("Datum","Q2A")],pub_s[,"WC_Q2A"]) 
colnames(indicator_plot) <- c("Date","Microdata","Published")
indicator_plot <- melt(indicator_plot, id="Date")  # convert to long format
g1 <- ggplot(data=indicator_plot,aes(x=Date, y=value, group=variable, colour=variable)) 
g1 <- g1 + geom_line()
g1 <- g1 + ylab("Indicator") + xlab("")
g1 <- g1 + theme(legend.title=element_blank())
g1 <- g1 + ggtitle("WC: Q2A") 
g1 <- g1 + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))
g1 <- g1 + theme(legend.title=element_blank()) 
g1 <- g1 + scale_x_date(labels = date_format("%Y"),breaks = date_breaks("year"))
g1 <- g1 + theme(legend.position="none")

indicator_plot <- cbind(GP.S[,c("Datum","Q2A")],pub_s[,"GP_Q2A"]) 
colnames(indicator_plot) <- c("Date","Microdata","Published")
indicator_plot <- melt(indicator_plot, id="Date")  # convert to long format
g2 <- ggplot(data=indicator_plot,aes(x=Date, y=value, group=variable, colour=variable)) 
g2 <- g2 + geom_line()
g2 <- g2 + ylab("") + xlab("")
g2 <- g2 + theme(legend.title=element_blank())
g2 <- g2 + ggtitle("GP: Q2A") 
g2 <- g2 + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))
g2 <- g2 + theme(legend.title=element_blank()) 
g2 <- g2 + scale_x_date(labels = date_format("%Y"),breaks = date_breaks("year"))
g2 <- g2 + theme(legend.position="none")

indicator_plot <- cbind(KZN.S[,c("Datum","Q2A")],pub_s[,"KZN_Q2A"]) 
colnames(indicator_plot) <- c("Date","Microdata","Published")
indicator_plot <- melt(indicator_plot, id="Date")  # convert to long format
g3 <- ggplot(data=indicator_plot,aes(x=Date, y=value, group=variable, colour=variable)) 
g3 <- g3 + geom_line()
g3 <- g3 + ylab("Indicator") + xlab("")
g3 <- g3 + theme(legend.title=element_blank())
g3 <- g3 + ggtitle("KZN: Q2A") 
g3 <- g3 + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))
g3 <- g3 + theme(legend.title=element_blank()) 
g3 <- g3 + scale_x_date(labels = date_format("%Y"),breaks = date_breaks("year"))
g3 <- g3 + theme(legend.position="bottom",plot.margin=unit(c(-0.5,0.4,0,0.4), "cm"))

indicator_plot <- cbind(WC.S[,c("Datum","Q2A")],GP.S[,c("Q2A")],KZN.S[,c("Q2A")]) 
colnames(indicator_plot) <- c("Date","WC","GP","KZN")
indicator_plot <- melt(indicator_plot, id="Date")  # convert to long format
g4 <- ggplot(data=indicator_plot,aes(x=Date, y=value, group=variable, colour=variable)) 
g4 <- g4 + geom_line()
g4 <- g4 + ylab("") + xlab("")
g4 <- g4 + theme(legend.title=element_blank())
g4 <- g4 + ggtitle("Regions: Q2A") 
g4 <- g4 + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))
g4 <- g4 + theme(legend.title=element_blank()) 
g4 <- g4 + theme(legend.position="bottom",plot.margin=unit(c(-0.5,0.4,0,0.4), "cm"))
g4 <- g4 + scale_x_date(labels = date_format("%Y"),breaks = date_breaks("year"))

library(gridExtra)
grid.arrange(g1, g2, g3, g4, ncol=2, nrow =2)
```

#Reference Series

##Reference Series
We still have to set up reference series

We can also try different weighting approaches



